%% —---------------------------------------------------
%% sbcreviews-en.tex
%% Template for SBC Reviews
%% (https://journals-sol.sbc.org.br/index.php/reviews/)
%%
%% A versão do template inclui algumas
%% modificações sobre o template original da REIC
%% (https://journals-sol.sbc.org.br/index.php/reic)
%%
%% Licença: CCBY 4.0
%% (https://creativecommons.org/licenses/by/4.0/)
%% Sociedade Brasileira de Computação.

\documentclass[english]{presets/sbcreviews-2025}%

\usepackage[misc,geometry]{ifsym}

%\usepackage{graphicx} % carregada dentro da classe
%\usepackage{fontspec} % carregada dentro da classe
%\usepackage{fontawesome} % carregada dentro da classe
%\usepackage{color} % carregada dentro da classe
%\usepackage{hyperref} % carregada dentro da classe

\usepackage{presets/aas_macros}
\usepackage[bottom]{footmisc}

\usepackage{tabularray}
% O pacote tabularray torna o uso dos pacotes abaixo desnecessarios.
%\usepackage{supertabular}
%\usepackage{multicol}
%\usepackage{multirow}

\usepackage{afterpage}
\usepackage{url}
\usepackage{pifont}

\setcitestyle{square}

\definecolor{engtitle}{rgb}{0.5,0.5,0.5}
\definecolor{orcidlogo}{rgb}{0.37,0.48,0.13}
\definecolor{unilogo}{rgb}{0.16, 0.26, 0.58}
\definecolor{maillogo}{rgb}{0.58, 0.16, 0.26}
\definecolor{darkblue}{rgb}{0.0,0.0,0.0}
\hypersetup{colorlinks,breaklinks,
  linkcolor=darkblue,urlcolor=darkblue,
anchorcolor=darkblue,citecolor=darkblue}
%\hypersetup{colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue}
% We disable hyperlinks by default with this line, to avoid the error "\pdfendlink ended up in different nesting level" while writing

\jid{ROCS}
\jtitle{SBC Computing Reviews, 2026, XX:1}
\issn{2966-3938}
\doi{10.5753/rocs.2026.XXXXXX}
\copyrightstatement{This work is licensed under a Creative Commons Attribution 4.0 International License}
\jyear{2026}

\category{Review}

\title[A Reproducible Benchmark of Apache IoTDB, InfluxDB, and TimescaleDB for IoT Workloads]{A Reproducible Benchmark of Apache IoTDB, InfluxDB, and TimescaleDB for IoT Workloads}

%THE ORCID IS MANDATORY FOR EACH AUTHOR IN ROCS
\author[Klein 2026]{
  \affil{\textbf{Luiz Fernando Klein}~\orcidlink{0009-0001-8949-3248}~\textcolor{blue}{\faEnvelope}~~[{Universidade Federal da Fronteira Sul - UFFS}~|\href{mailto: ...}{~{\textit{luizkleinfernando@gmail.com}}}~]}
}

\begin{document}

\begin{frontmatter}

  \maketitle

  \begin{mail}
    Computer Science Course, Universidade Federal da Fronteira Sul -- UFFS, Rodovia SC 484, Km 02, Chapecó, SC, 89815-899, Brazil.
  \end{mail}

  \begin{abstract-en}
    The rapid growth of Internet of Things deployments has generated unprecedented volumes of
    time-indexed sensor data, exposing the limitations of traditional databases and driving the
    adoption of specialized Time Series Databases (TSDBs). Despite growing interest, most published
    TSDB benchmarks lack sufficient scripts and configurations for independent replication.

    This paper presents a controlled, reproducible benchmark comparing Apache IoTDB~1.3.0,
    InfluxDB~2.7, and TimescaleDB (PostgreSQL~15) across nine IoT workload types at three
    data-volume scales spanning a hundredfold range in data size, using a fully scripted infrastructure
    built on Docker Compose and iot-benchmark.

    Results show that no single database dominates all workloads. IoTDB leads on write throughput
    (23.6\,M\,pts/s at large scale) and value-filter scaling (5.2$\times$ latency growth for
    a hundredfold increase in data). TimescaleDB provides sub-millisecond downsampling across all scales and the
    fastest latest-point lookups, backed by full SQL compatibility. InfluxDB offers a balanced
    middle ground with lower memory requirements. A key finding is that shared-resource benchmark
    environments produce misleading results: two conclusions from the initial runs were reversed
    by isolated reruns, underscoring that resource isolation is a prerequisite for valid comparisons.
  \end{abstract-en}

  \begin{keywords}
    Time Series Databases, Benchmarking, Apache IoTDB, InfluxDB, TimescaleDB, IoT
  \end{keywords}

  \begin{dates}
    % This information will be provided by the editor befeore publishing the paper
    \noindent{\sffamily\textbf{Received:}} DD Month YYYY~~~$\bullet$~~~
    {\sffamily\textbf{Accepted:}} DD Month YYYY~~~$\bullet$~~~
    {\sffamily\textbf{Published:}} DD Month YYYY

  \end{dates}

  %\begin{license}
  %Published under the Creative Commons Attribution 4.0 International Public License (CC BY 4.0)
  %\end{license}

\end{frontmatter}

% Chapters
\input{chapters/en_us/1_introduction}
\input{chapters/en_us/2_theoretical_foundation}
\input{chapters/en_us/3_related_works}
\input{chapters/en_us/4_methodology}
\input{chapters/en_us/5_experiments}
\input{chapters/en_us/6_conclusion}

\begin{declarations}

  \begin{contributions}
    L.F.K. designed the study, implemented the benchmark infrastructure, performed all experiments,
    analyzed the results, and wrote the manuscript.
  \end{contributions}

  \begin{interests}
    The authors declare that they have no competing interests.
  \end{interests}

  \begin{materials}
    The benchmark scripts, Docker Compose configurations, declarative build environment
    (Nix flake), and result datasets created and used in this study are publicly available at
    \url{https://github.com/LuizFerK/tcc2}.
  \end{materials}

  \begin{aitools}
    Claude Code (Anthropic, claude-sonnet-4-6) was used exclusively to assist in generating the Python scripts for benchmark runs and Matplotlib data analysis. The author independently performed all experimental design, result interpretation, and manuscript writing. All analytical conclusions and final decisions were made solely by the author, who takes full responsibility for the content.
  \end{aitools}

\end{declarations}

\bibliographystyle{presets/apalike-sol}
\bibliography{refs}

\end{document}
